User-User Collaborative Filtering
========================================================

Assumptions: The input data file is in the format "user item  weight" (tsv file with each entry on a new line & no header)
Other Inputs: user, number of recommendations, distance method, number of similar users to consider

```{r message=FALSE,echo=FALSE}

library("recommenderlab")

setwd("~/") #TODO: Pick the right path
userPreferencesDF <- read.csv(file="Abcam//AbcamSnowplow/redshiftetl/src/R/FinalWeightsCombined", #TODO: Parameterise
                              sep="\t",
                              colClasses=c("character","character","numeric"),
                              header=FALSE)
```

Sample input file is as follows:

```{r echo=FALSE}
head(userPreferencesDF)
```

Processing inputs...

```{r}
userColumn <- "user"
itemColumn <- "itemIDs"
weightColumn <- "weight"

colnames(userPreferencesDF)<- c(userColumn,itemColumn,weightColumn)
targetUserID <- "-3685863687683693081" #TODO Parameterise
```

Extract all those pages that the target user has viewed (irrespective of weights)

```{r}
viewedPages <- userPreferencesDF[which(userPreferencesDF[,userColumn]==targetUserID),itemColumn]
```

For this target user the pages viewed are:

```{r echo=FALSE}
print(viewedPages)
```

Fetch all users who have viewed at least one of these viewed pages

```{r}
usersWhoViewedSomeOfThosePages <- unique(userPreferencesDF[userPreferencesDF[,itemColumn] %in% viewedPages,
                                                           c(userColumn,itemColumn)])
```

This is how many such users exist in the data file

```{r echo=FALSE}
print(dim(usersWhoViewedSomeOfThosePages))
```

Now that we have the relevant users, find all the pages that they viewed

```{r}
pagesViewsBySimilarUsers <- aggregate(usersWhoViewedSomeOfThosePages[-1], #All the columns except the user column
                                      usersWhoViewedSomeOfThosePages[1],  #Group by user ID
                                      c)                                  #Concatenate all the item IDs per user
```

Now, per user count how many items are similar between this user & the target user

```{r}
pagesViewsBySimilarUsers$Count <- sapply(1:max(row(pagesViewsBySimilarUsers)), #Do this for all relevant users
                                         function(x){
                                           return (length(unlist(pagesViewsBySimilarUsers[x,2])));
                                           })
```

Post aggregation, this is how the processed data will look

```{r echo=FALSE}
head(pagesViewsBySimilarUsers)
```

Now we shall compute the rank of each user based on the number of pages in common with target user. We shall pick only the top N similar users.

```{r}
rankedUsersDF <- pagesViewsBySimilarUsers[order(pagesViewsBySimilarUsers$Count,decreasing=T),]
rankedUsersDF <- rankedUsersDF[-1,]
mostRelevantUsers <- rankedUsersDF[1:10,1] #TODO parameterise
```

These are the most relevant users for this target user:

```{r echo=FALSE}
print(mostRelevantUsers)
```

Now that we have the top N similar users, go back to the input file & extract all their page views alongwith the data for the target user

```{r}
pageViewDataOfTargetAndSimilarUsers <- userPreferencesDF[userPreferencesDF[,1] %in% c(targetUserID,mostRelevantUsers),]
```

The following code helps convert the data which is in a single-line-per-user-per-page-activity into a matrix of users & items with the weights in the appropriate cells. To keep data consistent across data structures, we will sort by user IDs

```{r}
tmpMatrix <- as(as(as(pageViewDataOfTargetAndSimilarUsers,"realRatingMatrix"),"dgCMatrix"),"matrix")
relevantUsersPageViewsMX <- cbind(row.names(tmpMatrix),tmpMatrix)
relevantUsersPageViewsMX <- relevantUsersPageViewsMX[order(relevantUsersPageViewsMX[,1]),]
itemsViewedByRelevantUsers <- colnames(relevantUsersPageViewsMX)
```

The transformation in data is shown below. 

Before : 

```{r echo=FALSE}
head(pageViewDataOfTargetAndSimilarUsers)
```

After : 

```{r echo=FALSE}
print(relevantUsersPageViewsMX[1:5, 1:5])
```

Identify all the items for which we have data as well as that subset which our relevant users have not viewed.

```{r}
allUniqueAbIDs <- unique(userPreferencesDF[,2])
abIDsNotViewedByRelevantUsers <- setdiff(unique(userPreferencesDF[,2]),itemsViewedByRelevantUsers)
```

Prepare a matrix which will be used to compute distances

```{r}
baseMX <- matrix(nrow=11, #TODO parameterise as N+1
                   ncol=length(abIDsNotViewedByRelevantUsers),
                   data=0)  #Initialise

colnames(baseMX) <- abIDsNotViewedByRelevantUsers
relevantUsersPrefsForAllItems <- cbind(c(targetUserID,mostRelevantUsers),baseMX)
relevantUsersPrefsForAllItems <- relevantUsersPrefsForAllItems[order(relevantUsersPrefsForAllItems[,1]),]
relevantUsersPrefsForAllItems <- cbind(relevantUsersPrefsForAllItems,relevantUsersPageViewsMX[,-1])
relevantUsersPrefsForAllItems <- rbind(relevantUsersPrefsForAllItems[relevantUsersPrefsForAllItems[,1] == targetUserID,], 
                                       relevantUsersPrefsForAllItems[relevantUsersPrefsForAllItems[,1] != targetUserID,])
colnames(relevantUsersPrefsForAllItems)[1] <- "userID"
userPageViewMatrixForDistanceComputation <- matrix(as.numeric(relevantUsersPrefsForAllItems[,-1]),
                                                   nrow=11) #TODO Parameterise as N+1
colnames(userPageViewMatrixForDistanceComputation) <- colnames(relevantUsersPrefsForAllItems)[-1]
```

Compute the distance using the appropriate method. This will result in an NxN matrix.

```{r}
distanceMX <- dist(userPageViewMatrixForDistanceComputation,
                     method="Manhattan", #TODO Parametrise
                     diag=T,
                     upper=T)

relevantUsersWeights <- as.matrix(distanceMX)[1,-1]
```

Here is the distance of users from each other. We are interested in the first row which is the distance from the target user.

```{r echo=FALSE}
print(relevantUsersWeights)
```

The weight for each user is inversely proportional to distance. Hence, we assign 1/distance, normalised to the closest user, as the weight. The weights are in range [0:1] and the most similar user has a weight 1.

```{r}
highestWeight <- min(relevantUsersWeights)
relevantUsersNormalWeights <- highestWeight/relevantUsersWeights
relevantUsersNormalWeightsMX <- matrix(relevantUsersNormalWeights,nrow=10) # TODO: parameterise
```

These are the weights for each user. 

```{r echo=FALSE}
print(relevantUsersNormalWeightsMX)
```

Apply the weights to the relevant users X pages matrix. Find the weights for each item. 

```{r}
weightedPagePreferenceMX <- userPageViewMatrixForDistanceComputation[-1,] * relevantUsersNormalWeightsMX[,1]
weightedPagePreferenceMX <- rbind(weightedPagePreferenceMX,colSums(weightedPagePreferenceMX))
pageRanks <- as.data.frame(weightedPagePreferenceMX[11,weightedPagePreferenceMX[11,]>0]) #TODO : parameterise
```

Order the pages by rank, remove the viewed pages & pick the top N as recommendations.

```{r}
pageRanks$AbIDs <- row.names(pageRanks)
pageRanks <- pageRanks[order(pageRanks[,1],decreasing=TRUE),]
newPagesToBeRecommended <- setdiff(pageRanks[,"AbIDs"],viewedPages)
```

Thus, the recommended pages are:

```{r echo=FALSE}
print(newPagesToBeRecommended[1:10]) #TODO Parameterise number of recommendations to provide
```